---
title: "Amazon Health Reviews Analysis"
author: "Nitish Neelagiri"
date: "April 16, 2016"
output: html_document
---

```{r}
set.seed(101)
library(stringr)
if(exists("HealthReviews") && is.data.frame(get('HealthReviews'))){
  
}else{
  HealthReviews <- read.csv("~/healthData.csv", stringsAsFactors=FALSE)
  helpfulnessData <- apply(data.frame(HealthReviews$helpful), 1, function(x){
  if(length(x)>0L){
  splitOutput <- strsplit(x, ',')[[1]]
  if (type.convert(strsplit(splitOutput[2], '\\]')[[1]][1]) == 0)
    helpfulness <- -1
  else{
  helpfulness <- type.convert(strsplit(splitOutput[1],'\\[')[[1]][2])/type.convert(strsplit(splitOutput[2],  '\\]')[[1]][1])}
  totalHelpfulness <- type.convert(strsplit(splitOutput[2], '\\]')[[1]][1])
  c(helpfulness,totalHelpfulness)
  }
})
  helpfulnessData <- t(helpfulnessData)
  helpfulnessData <- as.data.frame(helpfulnessData)
  names(helpfulnessData) <- c("HelpfulRatio", "TotalHelpful")
HealthReviews <- cbind(HealthReviews,helpfulnessData)
HealthReviews <- subset(HealthReviews, HealthReviews$TotalHelpful < 150 & HealthReviews$TotalHelpful > 15)
HealthReviews <- subset(HealthReviews, HealthReviews$HelpfulRatio >= 0 & HealthReviews$HelpfulRatio <= 1)

reviewTime <- apply(data.frame(HealthReviews$reviewTime),1, function(x){
    splitOutput1 <- strsplit(x, ',')
    splitOutput2 <- strsplit(splitOutput1[[1]][1], ' ')
    month <- type.convert(splitOutput2[[1]][1])
    day <- type.convert(splitOutput2[[1]][2])
    year <- type.convert(str_trim(splitOutput1[[1]][2]))
    c(day,month,year)
})
reviewTime <- t(reviewTime)
reviewTime <- as.data.frame(reviewTime)
names(reviewTime) <- c("Day", "Month", "Year")
HealthReviews <- cbind(HealthReviews, reviewTime)
HelpfulYears <- aggregate(TotalHelpful ~ Year, data = HealthReviews, FUN = sum)
ReviewFrequency <- data.frame(table(HealthReviews$Year))
names(ReviewFrequency) <- c("Year", "ReviewCount")
HelpfulPlot <- merge(HelpfulYears, ReviewFrequency)
plot(HelpfulPlot$TotalHelpful ~ HelpfulPlot$ReviewCount, main = "Number of helpfulness votes by reviews", xlab = "Review Count", ylab = "Helpfulness Votes Count")

reviewLength <- apply(data.frame(HealthReviews$reviewText), 1, function(x) nchar(x))
summary(reviewLength)
HealthReviews <- cbind(HealthReviews, reviewLength)
HealthReviews <- subset(HealthReviews, HealthReviews$reviewLength < 858)
dim(HealthReviews)


HealthReviewsTrain <- HealthReviews[HealthReviews$helpful >= 0,]
dim(HealthReviewsTrain)
rownames(HealthReviewsTrain) <- NULL
HealthReviewsTest <- HealthReviews[sample(nrow(HealthReviews),nrow(HealthReviews)-nrow(HealthReviewsTrain)),]
dim(HealthReviewsTest)
}
```

```{r}
head(HealthReviewsTrain)
class(HealthReviewsTrain$helpful)
reviewLength <- apply(data.frame(HealthReviewsTrain$reviewText), 1, function(x) nchar(x))
```

```{r}
hist(HealthReviewsTrain$overall)
```


Below boxplot describes the helpfulness based on the user rating. -1 indicates helpfulness value not available for a rating. This shows that most customers in the higher rating section such as (3,4,5) did not provide helpfulness for the product. The recommendation thus would be to improve efforts on data collection from customers who provide higher ratings.
```{r}
boxplot(out ~ HealthReviewsTrain$overall)

boxplot(reviewLengthScaled ~ HealthReviewsTrain$overall)

plot(reviewLength ~ HealthReviewsTrain$overall)
```

Review Text Analysis
```{r}
dfHelpfulness <- data.frame(HealthReviews$overall, HealthReviews$reviewText)
library(stringr)
yearCriterion <- apply(data.frame(HealthReviews$reviewTime), 1, function(x){
     type.convert(str_trim(strsplit(x, ',')[[1]][2], side="left"))
 })
dfHelpfulness <- data.frame(HealthReviews$helpful[k >= 2015], HealthReviews$reviewText[k >= 2015])
dim(dfHelpfulness)
dfHelpfulness <- data.frame(HealthReviews$asin[yearCriterion >= 2014], HealthReviews$helpful[yearCriterion >= 2014], HealthReviews$overall[yearCriterion >= 2014], HealthReviews$reviewText[yearCriterion >= 2014], HealthReviews$reviewerID[yearCriterion >= 2014], HealthReviews$summary[yearCriterion >= 2014])
dim(dfHelpfulness)
colnames(dfHelpfulness) <- c("helpful", "reviewText")

reviewSentiment <- apply(data.frame(HealthReviews$reviewText[HealthReviews$helpful >= 0]),1, function(x) getSentiment(annotateString(x)))


dfHelpfulness <- data.frame(HealthReviews$asin[yearCriterion >= 2014], HealthReviews$helpful[yearCriterion >= 2014], HealthReviews$overall[yearCriterion >= 2014], HealthReviews$reviewText[yearCriterion >= 2014], HealthReviews$reviewerID[yearCriterion >= 2014], HealthReviews$summary[yearCriterion >= 2014])
colnames(dfHelpfulness) <- c("asin", "helpful", "overall", "reviewText", "reviewerID", "summary")
dfHelpfulness$reviewText <- as.character(dfHelpfulness$reviewText)
dfHelpfulness$summary <- as.character(dfHelpfulness$summary)
dfHelpfulnessRated <- dfHelpfulness[dfHelpfulness$helpful >= 0,]
dim(dfHelpfulnessRated)
dfHelpfulnessNotRated <- dfHelpfulness[dfHelpfulness$helpful < 0,]
dim(dfHelpfulnessNotRated)

dfHelpfulnessRatedTrain <- dfHelpfulnessRated[sample(nrow(dfHelpfulnessRated[nchar(dfHelpfulnessRated$reviewText) < 500,]),5000),]
head(dfHelpfulnessRatedTrain)

dfHelpfulnessRatedTest <- dfHelpfulnessRated[sample(nrow(dfHelpfulnessRated),nrow(dfHelpfulnessRated)-nrow(dfHelpfulnessRatedTrain)),]
dim(dfHelpfulnessRatedTest)

reviewSentimentTrain <- apply(data.frame(dfHelpfulnessRatedTrain$reviewText),1, function(x){
  getSentiment(annotateString(x))
})


sentiment <- apply(data.frame(1:5000), 1, function(x){
    k <- count(if(is.null(reviewSentimentTrain[[x]][2])) 0 else reviewSentimentTrain[[x]][2])
    count <- 0
    out <- apply(data.frame(k), 1, function(y){
        if(is.na(y[1])) y[1] <- 0
        y[1] <- if (y[1] == 1) -1 else if (y[1] == 2) 0 else if (y[1] == 3) 1 else 0
        count = count + y[1]*y[2]
    })
    sum(data.frame(unlist(out))[1])
})

sentiment <- apply(data.frame(sentiment), 1, function(x){
    if (x < 0)
        "Negative"
    else
        if(x == 0)
            "Neutral"
    else
        "Positive"
})

predictorSet <- cbind(dfHelpfulnessRatedTrain$overall, sentiment, nchar(dfHelpfulnessRatedTrain$reviewText))
```
> totalSet[,1] <- apply(data.frame(totalSet[,1]), 1, function(x){
+     if(x > 0.7)
+         1
+     else
+         0
+ })
> trainTotalSet <- totalSet[sample(nrow(totalSet), 70*nrow(totalSet)/100),]
> testTotalSet <- totalSet[sample(nrow(totalSet),nrow(totalSet)-nrow(trainTotalSet)),]
> knn.pred = knn(trainTotalSet, testTotalSet, trainTotalSet[,1], k=1)
> table(knn.pred, testTotalSet[,1])